Introducing Distributed Dynamic Data-intensive (D3) Science: Understanding Applications and Infrastructure

### 2016-09-05 rebuttal.tex

\item \emph{``In this paper, we use locally distributed to indicate that data/compute resides on multiple nodes within one data center. Geographically distributed describes the distribution of data and/or compute across multiple data centers'' Given the above, I think it would be better in this definition to also separate out data creation location from compute location. Intro: Define precisely what is meant by ``dynamic,'' here as it is key to the paper but isn't well defined in Section 1 (see recommendation above to fix this by moving definitions from Section 3).}

\textbf{Response:}

\katznote{attention needed here re distributed and dynamic part of comment}

\nchnote{Partial response:} We have added a summary description of wha t we mean by distributed and dynamic at the start of Section 2 to reflect the fact that definition of the terminology arose and was refined by the discussions at the workshops, and the analysis of the application scenarios presented.

### 2016-09-05 3dpas_intro.tex
processing Yogesh's comments

\keywords{dynamic, distributed, data intensive, scientific applications}

\begin{abstract}
  A common feature across many science and engineering applications is the
  amount and diversity of data and computation that must be integrated to yield
  insights. Data sets are growing larger and becoming distributed; and their
  location, availability and properties are often time-dependent. Collectively, these
  characteristics give rise to dynamic distributed data intensive
  applications. While ``static'' data applications have received significant
  attention, the characteristics, requirements, and software systems for the
  analysis of large volumes of dynamic, distributed data, and data-intensive applications have
  received relatively less attention. This paper surveys several representative dynamic
  distributed data-intensive application scenarios, provides a common conceptual
  framework to understand them, and examines the infrastructure used in support
  of applications.

@@

The landscape of scientific and enterprise computing is being fundamentally
altered by prodigious volumes of data.  
%Occasionally data are often naturally
%distributed and have to be processed near their source, and sometimes data that
%are localized have to be distributed in order to be processed, Data can be naturally distributed because they are
%being created in a set of distributed locations, rather than in one place. 
Data can also be transported to be processed in multiple, distributed
locations, rather than being processed near their source, or be brought together for processing at
a particular location. These two orthogonal dimensions lead to a number of patterns of data
generation and processing being present in real applications,
either for reasons of performance or for collaborative analytics.  
Further, the variation of data production rates, data source, and destination makes 
temporal variations in data properties increasingly important.  
With an increase in the importance of distributed and temporal properties of 
data, systems and infrastructure aspects such as resource scheduling, 
data placement, and transfer decisions that
were statically determinable at small scales emerge as dynamic and important
considerations at large scales of data and distribution.

@@

\item Almost by definition, any attempt to capture the state-of-the-art of such
  a dynamic and emerging domain is bound to be fraught with limitations and
  issues of scoping.  A complete and rigorous survey is bound to be obsolete by
  the time it emerges; so given the rate of change, we asked what could be
  done. What we present is a partial analysis motivated by a set of existing and
  active D3 applications, their trials and tribulations, successes and
  solutions. We believe there is merit in capturing the state-of-play as is,
  deriving a research agenda and set of specific yet crosscutting issues, and
  presenting these to the community.

\item Distribution can occur on many levels.  In this paper, we use {\em locally
    distributed} to indicate that data/compute resides on multiple nodes within
  one data center.  {\em Geographically distributed} describes the distribution
  of data and/or compute across multiple data centers.  Also, our focus is
  primarily on infrastructure considerations at  large scale, as searching
  for common solutions and patterns at  large scale is likely to be more
  fruitful than at small scales where customized tools and optimization of
  low-level systems features is likely to return greater yield.

@@

Starting with the experience of the attendees of the 3DPAS workshops, we have
assembled a set of D3 ``applications'' that we describe in this section.  
%We realize that we have overloaded the term ``application'' here.

@@

Against this backdrop, we eschew a formal classification of application types,
however, to facilitate understanding we categorize applications into two
categories: the first is that of {\em traditional applications}, where
% traditional is  applications are {\em traditional applications}, in that 
a program (or set of programs) is developed by a user or project to try to
answer a science question (or set of related questions). 
A single standalone computer program, a distributed application, or a
complex workflow could fall under the category of a {\em traditional
application}.  Applications in this paper that are in this category are: NGS (next generation
sequencing) Analytics (\S\ref{bioSilvia}), CMB (cosmic microwave background)
(\S\ref{astroJulian}), Fusion (\S\ref{fusionScott}), Industrial Incident
Notification and Response (\S\ref{Kees}), MODIS (moderate resolution imaging
spectroradiometer) Data Processing (\S\ref{modisKeith}), and Distributed Network
Intrusion Detection (\S\ref{detectionJon}).

Another group of applications are {\em infrastructural applications}.
Applications of this type are collaborative, in that they depend on a context
that is often agreed upon by a community, most often in terms of how data is
stored by the community and how it is accessed, as well as, more often than not,
a weak consensus on the infrastructure to be used.  This allows different
sub-applications (that we again refer to as `applications') to focus on
different stages, such as generating data, or processing data.  Answering a
science question now may involve a set of applications that need to be run in
series, perhaps in different stages, that may be run by different groups that do
not frequently interact.  In this paper, examples of this type of application includes: ATLAS/WLCG (a
toroidal large hadron collider apparatus/worldwide large hadron collider
computing grid, \S\ref{WLCGSteve}), LSST (large synoptic survey telescope,
\S\ref{astroAdam}), SOA (service-oriented architecture) Astronomy
(\S\ref{astroSOAAdam}), Sensor Network Application (\S\ref{sensorSimon}),
Climate (\S\ref{climateDan}), Interactive Exploration of Environmental Data
(\S\ref{envJon}), and Power Grids (\S\ref{power}).

@@

In \S\ref{sec:distdyndata} we characterized dynamic and distributed
properties of the D3 Science applications. Our analysis was
qualitative by necessity, due to the fact that approaches to quantify
dynamism and distributed properties were often intimately tied to
application/scenario definition and specific usage modes.
Notwithstanding, to quantify better dynamic properties of an
application, we believe a term analogous to the use of Amdahl's
number~\cite{10.1109/ICDE.2000.839382} such as
\begin{itemize}
\item[] Num. of Compute Operations / Amount of Data
  (Consumed, Generated or Transformed)
\end{itemize}
  could be used to characterize `static'
data-intensive applications and systems:

%\begin{itemize}
%\item[] $\alpha$ = Num. of Compute Operations / Amount of Data
%  (Consumed, Generated or Transformed)
%\end{itemize}

Note that this term is scale-invariant, i.e., a value of 1 could
correspond to very small or very large values of both numerator and
denominator, and that it does not capture the time scales over
which data changes.

@@

Similarly, another term
\begin{itemize}
\item[] Amount of Data Transferred / Degree of
  Distribution
\end{itemize}
could be used to quantify the
``scale'' of distribution, namely the degree of distribution as a
function of the amount of data transferred, where the degree of
distribution can be defined as the number of distinct data locations:

@@

It has proven difficult to provide ``global'' quantitative values of
these ratios, as often the value of the numerators is
subject to how an application/scenario is defined, even more so for a
``class of applications'' as opposed to a specific application. We
believe, nonetheless, that these measures (which we refer to as $\alpha$ and $\beta$ respectively)
suggest a possible way to
quantify an application's dynamism and distribution and thereby find
possible common approaches solutions.  We plan to explore this in our
future work, and also hope others will consider building on these
initial concepts.

>>>>

\keywords{dynamic, distributed, data-intensive, scientific applications}

\begin{abstract}
  A common feature across many science and engineering applications is the
  amount and diversity of data and computation that must be integrated to yield
  insights. Data sets are growing larger and becoming distributed; and their
  location, availability and properties are often time-dependent. Collectively, these
  characteristics give rise to dynamic distributed data-intensive
  applications. While ``static'' data applications have received significant
  attention, the characteristics, requirements, and software systems for the
  analysis of large volumes of dynamic, distributed data, and data-intensive applications have
  received relatively less attention. This paper surveys several representative dynamic
  distributed data-intensive application scenarios, provides a common conceptual
  framework to understand them, and examines the infrastructure used in support
  of applications.

@@

The landscape of scientific and enterprise computing is being fundamentally
altered by prodigious volumes of data.  
%Occasionally data are often naturally
%distributed and have to be processed near their source, and sometimes data that
%are localized have to be distributed in order to be processed, Data can be naturally distributed because they are
%being created in a set of distributed locations, rather than in one place. 
This massive data generation can be naturally distributed because they are being created in a set of distributed
locations, rather than in one place. At the same time, data can also be
transported to be processed in multiple, distributed locations, rather than
being processed near their source, or be brought together for centrally
processing at a particular location. These two orthogonal dimensions lead to a
number of patterns of data generation and processing being present in real
applications, either for reasons of performance or for collaborative analytics.
Further, the variation of data production rates, data source, and destination
makes temporal variations in data properties increasingly important. With an
increase in the importance of distributed and temporal properties of data,
systems and infrastructure aspects such as resource scheduling, data placement,
and transfer decisions that were statically determinable at small scales emerge
as dynamic and important considerations at large scales of data and
distribution.

@@

\item Almost by definition, any attempt to capture the state-of-the-art of such
  a fast-changing and emerging domain is bound to be fraught with limitations 
  and issues of scoping.  A complete and rigorous survey is bound to be 
  obsolete by
  the time it emerges; so given the rate of change, we asked what could be
  done. What we present is a partial analysis motivated by a set of existing and
  active D3 applications, their trials and tribulations, successes and
  solutions. We believe there is merit in capturing the state-of-play as is,
  deriving a research agenda and set of specific yet crosscutting issues, and
  presenting these to the community.

\item Distribution can occur on many levels.  In this paper, we use {\em locally
    distributed} to indicate that data generation and storage, or computation 
	over it, occurs on multiple nodes within one data center.  
  {\em Geographically distributed} describes the distribution
  of data generation, storage and/or compute across multiple data centers.  
  Also, our focus is
  primarily on infrastructure considerations at  large scale, as searching
  for common solutions and patterns at  large scale is likely to be more
  fruitful than at small scales where customized tools and optimization of
  low-level systems features is likely to return greater yield.

 \item We use the term dynamism to refer to the spatio-temporal variability of 
 data (e.\,g.\ data rates and formats), applications (e.\,g., workflows that 
 depend on data values) and infrastructures (e.\,g.\ availability).

@@

Starting with the experience of the attendees of the Dynamic Distributed Data Programming Abstractions and Systems workshops (3DPAS), we have
assembled a set of D3 ``applications'' that we describe in this section.  
%We realize that we have overloaded the term ``application'' here.

@@

Against this backdrop, we eschew a formal classification of application types,
however, to facilitate understanding we categorize applications into two
categories: the first is that of {\em traditional applications}, where
% traditional is  applications are {\em traditional applications}, in that 
a program (or set of programs) is developed independently by a user or project to try to
answer a science question (or set of related questions). 
A single standalone computer program, a distributed application, or a
complex workflow could fall under the category of a {\em traditional
application}.  Applications in this paper that are in this category are: NGS (next generation
sequencing) Analytics (\S\ref{bioSilvia}), CMB (cosmic microwave background)
(\S\ref{astroJulian}), Fusion (\S\ref{fusionScott}), Industrial Incident
Notification and Response (\S\ref{Kees}), MODIS (moderate resolution imaging
spectroradiometer) Data Processing (\S\ref{modisKeith}), and Distributed Network
Intrusion Detection (\S\ref{detectionJon}).

Another group of applications are {\em infrastructural applications}.
Applications of this type are collaborative, in that they depend on a context
that is often agreed upon by a community, most often in terms of how data is
stored by the community and how it is accessed, as well as, more often than not,
a weak consensus on the infrastructure to be used.  This allows different
sub-applications (which themselves are full-fledged ``applications'', but with dependencies on the collaborative infrastructure) to focus on
different stages, such as generating data, or processing data.  Answering a
science question now may involve a set of applications that need to be run in
series, perhaps in different stages, that may be run by different groups that do
not frequently interact.  In this paper, examples of this type of application includes: ATLAS/WLCG (a
toroidal large hadron collider apparatus/worldwide large hadron collider
computing grid, \S\ref{WLCGSteve}), LSST (large synoptic survey telescope,
\S\ref{astroAdam}), SOA (service-oriented architecture) Astronomy
(\S\ref{astroSOAAdam}), Sensor Network Application (\S\ref{sensorSimon}),
Climate (\S\ref{climateDan}), Interactive Exploration of Environmental Data
(\S\ref{envJon}), and Power Grids (\S\ref{power}).

@@

In \S\ref{sec:distdyndata} we characterized dynamic and distributed
properties of the D3 Science applications. Our analysis was
qualitative by necessity, due to the fact that approaches to quantify
dynamism and distributed properties were often intimately tied to
application/scenario definition and specific usage modes.
Notwithstanding, to quantify better dynamic properties of an
application, we believe a term analogous to the use of Amdahl's
number~\cite{10.1109/ICDE.2000.839382}, such as $\alpha$ defined as Number\ of
Compute\ Operation / Amount\ Data (Consumed,\ Generated\ or\ Transformed)
% \begin{itemize}
% \item[] Num. of Compute Operations / Amount of Data
%   (Consumed, Generated or Transformed)
% \end{itemize}
could be used to characterize `static' data-intensive applications and systems.
%\begin{itemize}
%\item[] $\alpha$ = Num. of Compute Operations / Amount of Data
%  (Consumed, Generated or Transformed)
%\end{itemize}
Note that this term is scale-invariant, i.e., a value of 1 could
correspond to very small or very large values of both numerator and
denominator, and that it does not capture the time scales over
which data changes.

@@

Similarly, another term $\beta$ defined as Amount of Data Transferred / Degree of Distribution
% \begin{itemize}
% \item[] Amount of Data Transferred / Degree of
%   Distribution
% \end{itemize}
could be used to quantify the
``scale'' of distribution, namely the degree of distribution as a
function of the amount of data transferred, where the degree of
distribution can be defined as the number of distinct data locations.

@@

It has proven difficult to provide ``global'' quantitative values of
these ratios, as often the value of the numerators is
subject to how an application/scenario is defined, even more so for a
``class of applications'' as opposed to a specific application. We
believe, nonetheless, that these measures 
%(which we refer to as $\alpha$ and $\beta$ respectively)
suggest a possible way to quantify an application's dynamism and distribution 
and thereby find possible common approaches solutions.  We plan to explore this 
in our future work, and also hope others will consider building on these
initial concepts.
